---
title: "HW7_SDS315"
author: "Anh Doan (atd2354)"
output:   
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(mosaic)
library(tidyverse)
library(MatchIt)

armfold <- read.csv("armfold.csv")
turnout <- read.csv("turnout.csv")
```

# **Problem 1: Armfolding**

### **Part A**

```{r}
gender <- table(armfold$Sex)

male <- armfold[armfold$Sex == "Male", ]
male_left <- mean(male$LonR_fold == 1)

female <- armfold[armfold$Sex == "Female", ]
female_left <- mean(female$LonR_fold == 1)

cat(
  "The Number of Male Students: ", gender["Male"], "\n",
  "The Number of Female Students: ", gender["Female"], "\n",
  "The Proportion of Males Who Folded Their Left Arm On Top: ", round(male_left, 4), "\n",
  "The Proportion of Females Who Folded Their Left Arm On Top: ", round(female_left, 4), "\n",
  sep = ""
)
```

### **Part B**

```{r}
cat("The observed difference in proportions between the two groups is", diff_prop <- round(diffprop(LonR_fold ~Sex, data = armfold), 4), "\n")
```

### **Part C**

##### **- R Built-in Functions**

```{r}
#Set seed for reproducibility
set.seed(123)

#Bootstrap the difference in proportions between the two groups
prop_left <- do(10000)*diffprop(LonR_fold ~Sex, data = resample(armfold))

#Construct the confidence interval
confint(prop_left, level=0.95)
```

##### **- "Hand-Calculated" Version**

<center>
  ![**Standard Error Formula For Difference in Proportions**](https://lh6.googleusercontent.com/proxy/DHp1FIkdd2SdR1ZPG5PYMTmvio1KgTZP_uCJHFYTJa_UYv-Jpy0uPDrM8WQ0r6bEVwiwBOjDrgwoQ1fdbtM3hDiNY3tf4A5AJaFY1w)
</center>

- $\hat{p_{1}}$ = `r male_left` (Sample Proportion in the Male Group)
- $\hat{p_{2}}$ = `r female_left` (Sample Proportion in the Female Group)
- $n_{1}$ = **106** (Sample Size of the Male Group)
- $n_{2}$ = **111** (Sample Size of the Female Group) 
- $z^{*}$ = **1.96** (The Critical Value For a 95% Confidence Interval)

```{r, echo = TRUE}
p1 = male_left
p2 = female_left
n1 = 106
n2 = 111
z = 1.96

#Find the Standard Error
SE <- sqrt((p1 * (1 - p1) / n1) + (p2 * (1 - p2) / n2))

#Construct The Confidence Interval
diff_prop <- p1 - p2
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE
```

```{r}
cat(
  "Standard error: ", SE, "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n",
  sep = ""
)
```


### **Part D**

If we were to repeat this sampling process many times, the true difference in proportions between males and females would fall within the confidence interval (-0.08, 0.18) at least 95% of the time. 

### **Part E**

The **standard error** represents the statistical fluctuation of the typical magnitude of error.

### **Part F**

The term **sampling distribution** refers to the distribution of the difference in proportions between males and females we expect to see for a statistical summary under repeated sampling. The test statistic (difference in proportions) and proportions of males and females ($\hat{p_{1}}$) and ($\hat{p_{2}}$) vary from sample to sample, while the sample size remains fixed. 

### **Part G**

The mathematical result mathematical that justifies using a normal distribution to approximate the sampling distribution of the difference in sample proportions is the **Central Limit Theorem (CLT)**. It states that, as the sample size increases, the sampling distribution of the mean is approximately normal, regardless of the shape of the original population distribution.

### **Part H**

The 95% confidence interval includes 0, suggesting that there is a possibility that there is no difference between the two groups. Therefore, the claim "there’s no sex difference in arm folding” is **not supported by the data.**

### **Part I**

The confidence interval would be **different** if we repeat this experiment many times with different random samples of university students due to random sampling variability and sampling error. The collection of many confidence intervals should cover the corresponding true value at least X% of the time, if we were to construct X% confidence intervals. 

# **Problem 2: Get out the vote**

## **Part A**
 
```{r}
prop_call <- mean(turnout$voted1998[turnout$GOTV_call == 1])
prop_no_call <- mean(turnout$voted1998[turnout$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout$GOTV_call == 1)
n_no_gotv <- sum(turnout$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of those receiving a GOTV call who voted in 1998 is ", round(prop_call, 4), "\n",
  "The proportion of those not receiving a GOTV call who voted in 1998 is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Conclusion: Based on the 95% confidence interval, recipients of the GOTV call were between 14.32% and 26.38% more likely to vote in 1998 compared to those who did not receive the call.",
  sep = ""
)
```

## **Part B**
### **Voting Status 1996**
##### **Voting Status 1996 vs GOTV Call**

```{r}
prop_call <- mean(turnout$voted1996[turnout$GOTV_call == 1])
prop_no_call <- mean(turnout$voted1996[turnout$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout$GOTV_call == 1)
n_no_gotv <- sum(turnout$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of those receiving a GOTV call who voted in 1996 is ", round(prop_call, 4), "\n",
  "The proportion of those not receiving a GOTV call who voted in 1996 is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, recipients of the GOTV call were between 12.45% and 23.9% more likely to have voted in 1996 compared to those who did not receive the call, suggesting an association between receiving a GOTV call and 1996 voting status.",
  sep = ""
)
```

##### **Voting Status 1996 vs Voting Status 1998**

```{r}
prop_voted <- mean(turnout$voted1996[turnout$voted1998 == 1])
prop_no_voted <- mean(turnout$voted1996[turnout$voted1998 == 0])

#Sample Sizes
n_voted <- sum(turnout$voted1998 == 1)
n_no_voted <- sum(turnout$voted1998 == 0)

#Standard error formula
SE <- sqrt((prop_voted * (1 - prop_voted) / n_voted) + (prop_no_voted * (1 - prop_no_voted) / n_no_voted))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_voted - prop_no_voted
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of those who voted in 1998 who voted in 1996 is ", round(prop_voted, 4), "\n",
  "The proportion of those who did not vote in 1998 who voted in 1996 is ", round(prop_no_voted, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, 1998 voters were between 39.57% and 42.97% more likely to have voted in 1996 compared to those who did not vote in 1998, suggesting an association between 1998 voting status and 1996 voting status.",
  sep = ""
)
```

##### **Interpretation**

Therefore, **"voted1996"** is a confounder because it is associated with both receiving a GOTV call and voting in 1998, preventing the observed difference in Part A from representing the true causal effect of the GOTV call on the likelihood that a person voted in 1998. 

### **Age**
##### **Age vs GOTV Call**

```{r}
mean_age_call <- mean(turnout$AGE[turnout$GOTV_call == 1])
mean_age_no_call <- mean(turnout$AGE[turnout$GOTV_call == 0])

#Difference of Means
diff_call <- mean_age_call - mean_age_no_call

# Standard deviations
sd_age_call <- sd(turnout$AGE[turnout$GOTV_call == 1])
sd_age_no_call <- sd(turnout$AGE[turnout$GOTV_call == 0])

# Sample sizes
n_call <- sum(turnout$GOTV_call == 1)
n_no_call <- sum(turnout$GOTV_call == 0)

# Standard error formula
SE <- sqrt((sd_age_call^2 / n_call) + (sd_age_no_call^2 / n_no_call))

# Construct the Confidence Interval
z <- 1.96
lower <- diff_call - z * SE
upper <- diff_call + z * SE

#Output results
cat(
  "The average age of those who received a GOTV call is ", round(mean_age_call, 2), " years,\n",
  "The average age of those who did not receive a cal is ", round(mean_age_no_call, 2), " years.\n",
  "Difference of Mean: ", round(diff_call, 2), " years\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, recipients of the GOTV call were between 6.38 and 11.38 years older than those who did not receive the call, suggesting an association between receiving a GOTV call and age.",
  sep = ""
)
```

##### **Age vs Voting Status 1998**

```{r}
mean_age_voted <- mean(turnout$AGE[turnout$voted1998 == 1])
mean_age_no_voted <- mean(turnout$AGE[turnout$voted1998 == 0])

#Difference of Means
diff_voted <- mean_age_voted - mean_age_no_voted

# Standard deviations
sd_age_voted <- sd(turnout$AGE[turnout$voted1998 == 1])
sd_age_no_voted <- sd(turnout$AGE[turnout$voted1998 == 0])

# Sample sizes
n_voted <- sum(turnout$voted1998 == 1)
n_no_voted <- sum(turnout$voted1998 == 0)

# Standard error formula
SE <- sqrt((sd_age_voted^2 / n_voted) + (sd_age_no_voted^2 / n_no_voted))

# Construct the Confidence Interval
z <- 1.96
lower <- diff_voted - z * SE
upper <- diff_voted + z * SE

#Output results
cat(
  "The average age of those who received a GOTV call is ", round(mean_age_voted, 2), " years,\n",
  "The average age of those who did not receive a cal is ", round(mean_age_no_voted, 2), " years.\n",
  "Difference of Mean: ", round(diff_voted, 2), " years\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, 1998 voters were between 9.82 and 11.18 years older than those who did not vote in 1998, suggesting an association between 1998 voting status and age.",
  sep = ""
)
```

##### **Interpretation**

Therefore, **"AGE"** is a confounder because it is associated with both receiving a GOTV call and voting in 1998, preventing the observed difference in Part A from representing the true causal effect of the GOTV call on the likelihood that a person voted in 1998. 

### **MAJORPTY**
##### **MAJORPTY vs GOTV Call**

```{r}
prop_call <- mean(turnout$MAJORPTY[turnout$GOTV_call == 1])
prop_no_call <- mean(turnout$MAJORPTY[turnout$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout$GOTV_call == 1)
n_no_gotv <- sum(turnout$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of GOTV call recipients registered with a major U.S. political party is ", round(prop_call, 4), "\n",
  "The proportion of GOTV call recipients not registered with a major U.S. political party is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, recipients of the GOTV call were between 0.64% and 10.73% more likely to be registered with a major U.S. political party compared to those who did not receive the call, suggesting an association between receiving a GOTV call and major party affiliation.",
  sep = ""
)
```

##### **MAJORPTY vs Voting Status 1998**

```{r}
prop_voted <- mean(turnout$MAJORPTY[turnout$voted1998 == 1])
prop_no_voted <- mean(turnout$MAJORPTY[turnout$voted1998 == 0])

#Sample Sizes
n_voted <- sum(turnout$voted1998 == 1)
n_no_voted <- sum(turnout$voted1998 == 0)

#Standard error formula
SE <- sqrt((prop_voted * (1 - prop_voted) / n_voted) + (prop_no_voted * (1 - prop_no_voted) / n_no_voted))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_voted - prop_no_voted
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of GOTV call recipients registered with a major U.S. political party is ", round(prop_voted, 4), "\n",
  "The proportion of GOTV call recipients not registered with a major U.S. political party is ", round(prop_no_voted, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Interpretation: Based on the 95% confidence interval, 1998 voters were between 8.52% and 11.75% more likely to be registered with a major U.S. political party compared to those who did not vote in 1998, suggesting an association between 1998 voting status and major party affiliation.",
  sep = ""
)
```

##### **Interpretation**

Therefore, **"MAJORPTY"** is a confounder because it is associated with both receiving a GOTV call and voting in 1998, preventing the observed difference in Part A from representing the true causal effect of the GOTV call on the likelihood that a person voted in 1998. 

## **Part C**

```{r}
# Construct a data set with GOTV_call as our treatment variable, and with voted1996, AGE, and MAJORPTY as “matching” or “balancing” variables. 
turnout_match <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, 
                     data = turnout, 
                     ratio = 5)

# Extract the matched data 
turnout_matched <- match.data(turnout_match)
```

##### **Confounder Check: Voting Status in 1996**

```{r}
prop_call <- mean(turnout_matched$voted1996[turnout_matched$GOTV_call == 1])
prop_no_call <- mean(turnout_matched$voted1996[turnout_matched$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout_matched$GOTV_call == 1)
n_no_gotv <- sum(turnout_matched$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of those receiving a GOTV call who voted in 1996 is ", round(prop_call, 4), "\n",
  "The proportion of those not receiving a GOTV call who voted in 1996 is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Based on summary statistics, the matched data set is balanced with respect to the voted1996 confounder",
  sep = ""
)
```

##### **Confounder Check: Age**

```{r}
mean_age_call <- mean(turnout_matched$AGE[turnout_matched$GOTV_call == 1])
mean_age_no_call <- mean(turnout_matched$AGE[turnout_matched$GOTV_call == 0])

#Difference of Means
diff_call <- mean_age_call - mean_age_no_call

# Standard deviations
sd_age_call <- sd(turnout_matched$AGE[turnout_matched$GOTV_call == 1])
sd_age_no_call <- sd(turnout_matched$AGE[turnout_matched$GOTV_call == 0])

# Sample sizes
n_call <- sum(turnout$GOTV_call == 1)
n_no_call <- sum(turnout$GOTV_call == 0)

# Standard error formula
SE <- sqrt((sd_age_call^2 / n_call) + (sd_age_no_call^2 / n_no_call))

# Construct the Confidence Interval
z <- 1.96
lower <- diff_call - z * SE
upper <- diff_call + z * SE

#Output results
cat(
  "The average age of those who received a GOTV call is ", round(mean_age_call, 2), " years,\n",
  "The average age of those who did not receive a cal is ", round(mean_age_no_call, 2), " years.\n",
  "Difference of Mean: ", round(diff_call, 2), " years\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Based on summary statistics, the matched data set is balanced with respect to the AGE confounder",
  sep = ""
)
```

##### **Confounder Check: MAJORPTY**

```{r}
prop_call <- mean(turnout_matched$MAJORPTY[turnout_matched$GOTV_call == 1])
prop_no_call <- mean(turnout_matched$MAJORPTY[turnout_matched$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout_matched$GOTV_call == 1)
n_no_gotv <- sum(turnout_matched$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of GOTV call recipients registered with a major U.S. political party is ", round(prop_call, 4), "\n",
  "The proportion of GOTV call recipients not registered with a major U.S. political party is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Based on summary statistics, the matched dataset is balanced with respect to the MAJORPTY confounder",
  sep = ""
)
```

##### **Analysis** 

```{r}
prop_call <- mean(turnout_matched$voted1998[turnout_matched$GOTV_call == 1])
prop_no_call <- mean(turnout_matched$voted1998[turnout_matched$GOTV_call == 0])

#Sample Sizes
n_gotv <- sum(turnout_matched$GOTV_call == 1)
n_no_gotv <- sum(turnout_matched$GOTV_call == 0)

#Standard error formula
SE <- sqrt((prop_call * (1 - prop_call) / n_gotv) + (prop_no_call * (1 - prop_no_call) / n_no_gotv))

#z* for confidence interval
z <- 1.96

#Construct The Confidence Interval
diff_prop <- prop_call - prop_no_call
lower <- diff_prop - z * SE
upper <- diff_prop + z * SE

#Output Results
cat(
  "The proportion of those receiving a GOTV call who voted in 1998 is ", round(prop_call, 4), "\n",
  "The proportion of those not receiving a GOTV call who voted in 1998 is ", round(prop_no_call, 4), "\n",
  "Difference in Proportions: ", round(diff_prop, 4), "\n",
  "Standard Error: ", round(SE, 4), "\n",
  "95% Confidence Interval: [", round(lower, 4), ", ", round(upper, 4), "]\n\n",
  "Conclusion: Based on the 95% confidence interval, recipients of the GOTV call were between 1.29% and 14.42% more likely to vote in 1998 compared to those who did not receive the call, suggesting a positive effect of GOTV Call on 1998 voter turnout after matching",
  sep = ""
)
```

